Best Practices in Mixture Modeling using Free Open Source Software

### 2022-08-01 manuscript.Rmd
add rstudio project, move open science paragraph, fix shoesize

```{r load_packages, include = FALSE}
library("papaja")
#bibliography      : ["r-references.bib"]
library(worcs)
```

@@

Throughout this paper, examples make use of free, open source software for latent class analysis in R. 
As the open science movement is gaining momentum,
researchers are becoming increasingly aware
how important it is that analyses can be reproduced and audited.
In line with open science principles, one of the suggested reporting standards relates to reproducible code.
In this context, it is important to note that user-friendly methods for estimating latent class analyses have predominantly been available in commercial software packages (e.g., *Mplus* and *Latent GOLD*).
A potential downside of commercial software is that it restricts the ability to reproduce analyses to license holders,
and prevents auditing research because the underlying source code is proprietary.
To overcome these limitations, the present paper introduces new user-friendly functions in the `tidySEM` R-package that can be used to estimate a wide range of latent class analysis models using the free, open-source R-package `OpenMx`.
The reporting guidelines described in this paper are adopted in `tidySEM` by default.
The `tidySEM` R-package thus makes advanced mixture modeling based on best practices widely accessible,
and facilitates the adoption of the estimation and reporting guidelines described in this paper.

@@

```{r shoedens, echo = FALSE, fig.cap="Kernel density plot of shoe sizes."}
library(tidySEM)
library(ggplot2)
set.seed(1)
n = 100
C <- sample(c("Man", "Woman"), n, replace = TRUE)
means <- c(Man = 9, Woman = 7)
X <- rnorm(n, mean = means[C], sd = 1)
X <- round(X*2)/2
obsparam <- tapply(X, C, mean)
est <- mx_profiles(data.frame(X), 2)
estparam <- table_results(est, c("label", "est", "matrix"))
estparam <- estparam$est[estparam$matrix == "M"]
props <- class_prob(est, "individual")$individual
props <- apply(props, 1, which.max)

ggplot(data.frame(Shoesize = X), aes(x = Shoesize)) + geom_density() + theme_bw()
```

@@

>>>>

```{r load_packages, include = FALSE}
library("papaja")
#bibliography      : ["r-references.bib"]
library(worcs)
run_everything = FALSE
```

@@

Throughout this paper, examples make use of free, open source software for latent class analysis in R.

@@

```{r echo = FALSE}
library(tidySEM)
library(ggplot2)
set.seed(1)
n = 100
C <- sample(c("Man", "Woman"), n, replace = TRUE)
means <- c(Man = 9, Woman = 7)
X <- rnorm(n, mean = means[C], sd = 1)
X <- round(X*2)/2
obsparam <- tapply(X, C, mean)
```

```{r echo = FALSE, eval = run_everything}
est <- mx_profiles(data.frame(X), 2)
estparam <- table_results(est, c("label", "est", "matrix"))
estparam <- as.numeric(estparam$est[estparam$matrix == "M"])
props <- class_prob(est, "individual")$individual[ , "predicted"]
dput(props, "props.txt")
dput(estparam, "estparam.txt")

p <- ggplot(data.frame(Shoesize = X), aes(x = Shoesize)) + geom_density() + theme_bw()
ggsave("shoedens.png", p, device = "png")
```

```{r shoedens, echo = FALSE, fig.cap="Kernel density plot of shoe sizes."}
props <- eval(parse("props.txt"))
estparam <- eval(parse("estparam.txt"))
knitr::include_graphics("shoedens.png")
```

@@

<!-- CJ: I think this belongs here -->
As the open science movement is gaining momentum,
researchers are becoming increasingly aware
how important it is that analyses can be reproduced and audited.
In line with open science principles, one of the suggested reporting standards relates to reproducible code.
In this context, it is important to note that user-friendly methods for estimating latent class analyses have predominantly been available in commercial software packages (e.g., *Mplus* and *Latent GOLD*).
A potential downside of commercial software is that it restricts the ability to reproduce analyses to license holders,
and prevents auditing research because the underlying source code is proprietary.
To overcome these limitations, the present paper introduces new user-friendly functions in the `tidySEM` R-package that can be used to estimate a wide range of latent class analysis models using the free, open-source R-package `OpenMx`.
The reporting guidelines described in this paper are adopted in `tidySEM` by default.
The `tidySEM` R-package thus makes advanced mixture modeling based on best practices widely accessible,
and facilitates the adoption of the estimation and reporting guidelines described in this paper.
